---
title: "PH 1831"
output:
  html_document:
    df_print: paged
date: "2022-12-03"
---

```{r setup, include=FALSE}
library(tidyverse)
library(survival)
```

# Variables Key
futime   = number of days between registration and the earlier of death and censoring
status   = 0=censored; 1=died.
age      = age in years
drug     = 0=placebo, 1=treatment
sex      = 0=male, 1=female
ascites  = presence of ascites: 0=no 1=yes
hepato   = presence of hepatomegaly 0=no 1=yes
spiders  = presence of spiders 0=no 1=yes
edema    = presence of edema 0=no edema and no diuretic therapy for edema;
          .5 = edema present without diuretics, or edema resolved by diuretics;
           1 = edema despite diuretic therapy
bili     = serum bilirubin in mg/dl
chol     = serum cholesterol in mg/dl
albumin  = albumin in gm/dl
copper   = urine copper in ug/day
alk_phos = alkaline phosphatase in U/liter
sgot     = SGOT in U/ml
trig     = triglicerides in mg/dl
platelet = platelets per cubic ml/1000
protime  = prothrombin time in seconds
stage    = histologic stage of disease


```{r}
data<- read.csv("/Users/shefalipatel/Documents/Fall 2022/Survival Analysis/Project/PBC276.csv")
(data)
```

```{r}
#determine variable type
str(data)
#summary stats for numeric variables
data_numeric <- data[,-c(3,4,6,7,8,9,10)]
summary(data_numeric)
#frequency table for binary variables 
data_binary <- data[,c(3,4,6,7,8,9,10)]
data_binary
apply((data_binary), 2, table)
```


```{r}
#subset by drug = 1
data_1 <- subset(data, drug == 1)
data_1
# subset by drug = 0 
data_0 <- subset(data, drug == 0)
data_0
```

```{r}
#univariate split by drug 
bivariate <- c(3,4,6,7,8,9,10)
summary(data_0[,-bivariate])
apply((data_0[,bivariate]), 2, table)
summary(data_1[,-bivariate])
apply((data_1[,bivariate]), 2, table)
```

```{r}
colnames(data)
```

```{r}
#make intial model
res.cox <- coxph(Surv(futime, status) ~ factor(drug) + age + factor(sex) + factor(ascites) + factor(hepato) + factor(spiders) + factor(edema) + bili
                 + chol + albumin + copper + alk_phos + sgot + trig + platelet + protime + stage, data =  data)
res.cox
summary(res.cox)
```

```{r}
plot(survfit(res.cox), ylab="Probability of Survival", xlab="Days", 
     col=c("red", "black", "black"))
```


```{r}
#stepwise
step(res.cox)
```

```{r}
res.cox.selected <- coxph(formula = Surv(futime, status) ~ age + factor(edema) + 
    bili + albumin + copper + sgot + protime + stage, data = data)
res.cox.selected
summary(res.cox.selected)

```

```{r}
plot(survfit(res.cox.selected), ylab="Probability of Survival", xlab="Days", 
     col=c("red", "black", "black"))
```

```{r}
median(data$copper)
median(data$sgot)
median(data$protime)
data$copper_split <- ifelse(data$copper > 74, 1, 0)
data$sgot_split <- ifelse(data$sgot > 116.625, 1, 0)
data$protime_split <- ifelse(data$protime > 10.6, 1, 0)
```


```{r}
library(survminer)
ggcoxdiagnostics(res.cox.selected, type = c("martingale"))
ggcoxfunctional(Surv(futime, status) ~ age  + 
    bili + albumin + copper + sgot + protime + stage, data = data, fit=res.cox.selected)
#age and bili and albumin potential linearity 
#copper, sgot, protime show non-linearity, discretized version needed 
#edema factor 1 not able to be plotted using ggcoxfunctional command
```
```{r}
res.cox.transformed <- coxph(formula = Surv(futime, status) ~ age + factor(edema) + 
    log(bili) + albumin + factor(copper_split) + factor(sgot_split) + factor(protime_split) + stage, data = data)
res.cox.transformed
ggcoxdiagnostics(res.cox.transformed, type = c("martingale"))
ggcoxfunctional(Surv(futime, status) ~ age + 
    log(bili) + albumin + stage, data = data, fit=res.cox.transformed)

```


```{r}
test.ph <- cox.zph(res.cox.transformed)
test.ph
ggcoxdiagnostics(res.cox.transformed, type = c("schoenfeld"))

```


```{r}
ggcoxdiagnostics(res.cox.transformed, type = c("deviance"))

```


```{r}
ggcoxdiagnostics(res.cox.transformed, type = c("dfbeta"))

```


```{r}
data$resid_mart <- residuals(res.cox.transformed, type = "martingale")
## Cox-Snell residuals
data$resid_coxsnell <- -(data$resid_mart - data$status)


## Fit model on Cox-Snell residuals (Approximately Expo(1) distributed under correct model)
fit_coxsnell <- coxph(formula = Surv(resid_coxsnell, status) ~ 1,
                      data    = data,
                      ties    = c("efron","breslow","exact")[1])

## Nelson-Aalen estimator for baseline hazard (all covariates zero)
df_base_haz <- basehaz(fit_coxsnell, centered = FALSE)

```


```{r}
ggplot(data = df_base_haz, mapping = aes(x = time, y = hazard)) +
    geom_point() +
    scale_x_continuous(limit = c(0,3.5)) +
    scale_y_continuous(limit = c(0,3.5)) +
    labs(x = "Cox-Snell residuals as pseudo observed times",
         y = "Estimated cumulative hazard at pseudo observed times") +
    theme_bw() + theme(legend.key = element_blank())

```

```{r}
plot(survfit(res.cox.transformed), ylab="Probability of Survival", xlab="Days", 
     col=c("red", "black", "black"))
```

```{r}
summary(res.cox.transformed)
```

